---
title: 'Blue Paper 10: habitats'
resource_files:
- data/EEZ_land.shx
- data/EEZ_land.prj
- data/EEZ_land.dbf
- data/EEZ_ocean.cpg
- data/EEZ_ocean.dbf
- data/EEZ_ocean.prj
- data/EEZ_ocean.qpj
- data/EEZ_ocean.shx
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    source_code: embed
    theme: spacelab
    vertical_layout: scroll
---

```{r global, include=FALSE}
library(flexdashboard)
library(datasets)
library(tidyverse)
library(stringr)
library(mapview)
library(maptools)
library(raster)
library(rgeos)
library(DT)
library(vegan)
library(plotly)
library(shiny)
library(ggthemes)
library(leaflet)
world <- shapefile("data/EEZ_ocean.shp", verbose = FALSE)
habitats <- read.csv("data/habitats.csv")

cntry <- habitats %>%
  group_by(Zone, Habitat) %>%
  mutate(total_area = sum(Area)) %>%
  group_by(Country, Zone, Habitat) %>% 
  summarise(rel_area = (Area/total_area)*100)


```

Dataset
=======================================================================

Column {data-height=700}
-----------------------------------------------------------------------

### All habitat data were intersected with Exclusive Economical Zone layer to calculate the area in ```km^2```, then were divided in coastal habitats and oceanic habitats. 
```{r}
datatable(
  habitats %>%
    select(Country, Zone, Habitat, Area),
  filter = 'top',
  rownames = FALSE
) %>%
  formatRound("Area", digits = 1)
```

Column {data-height=700}
-----------------------------------------------------------------------

### Richnes values are totals per each zone: Oceanic and Coastal, same for Shannon H: use filter or search to show different Countries and the next button below to scroll down the rows.
```{r Datatable diversity}
habitats %>%
  add_count(Country, Zone) %>%
  group_by(Country, Zone) %>%
  mutate(H = diversity(Area, index = "shannon", 2))  %>%
  select(Country, Zone, Richness = n, Shannon_H = H) %>%
  unique(.) -> rich

datatable(rich,
          filter = 'top',
          rownames = FALSE) %>%
  formatRound("Shannon_H", digits = 1)
```


Habitats
=======================================================================


Column {.sidebar}
-----------------------------------------------------------------------

This allow to select specific habitat to show their habitat area contribution

```{r}
selectInput(
      inputId = "Habitat",
      label = NULL,
      choices = sort(unique(cntry$Habitat))
      )
```

These % are calculated over the total area of each habitat, here we represent the contribution each country has to the total in %. 


Column
-----------------------------------------------------------------------

### Top 20 countries for habitat contribution in %

```{r}

dfInput <- reactive({
        cntry %>% ungroup() %>% filter(Habitat == input$Habitat)
    })
    
renderPlot({
        df1 <- dfInput() %>% arrange(Country, rel_area) %>% top_n(20)
        ggplot(df1) +
            aes(x = reorder(Country, rel_area), y = rel_area) +
            geom_col(fill = "black") +
            labs(x = "", y = "Relative % to total habitat area") +
            theme_economist_white() +
            theme(axis.text.x = element_text(angle = 90))+
            coord_flip()
    })

```



Countries
=======================================================================


Column {.sidebar}
-----------------------------------------------------------------------

This allow to select a specific Country and Zone to show their habitat area contribution in %

```{r}
selectInput(
      inputId = "Country",
      label = NULL,
      choices = unique(cntry$Country)
      )

selectInput(
      inputId = "Zone",
      label = NULL,
      choices = unique(cntry$Zone)
      )
```

These % are calculated over the total area of each habitat, here we represent the contribution each country has to the total in %. 

Column
-----------------------------------------------------------------------

### Top 20 countries for habitat area

```{r}
dfInput2 <- reactive({
        cntry %>% filter(Country == input$Country & Zone == input$Zone)
    })
    
renderPlot({
        df2 <- dfInput2() %>% arrange(Habitat, rel_area) 
        ggplot(df2) +
            aes(x = reorder(Habitat, rel_area), y = rel_area) +
            geom_col(fill = "black") +
            labs(x = "", y = "Relative % to total habitat area") +
            theme_economist_white() +
            theme(axis.text.x = element_text(angle = 90))+
            coord_flip()
    })
```


Coastal map
=======================================================================

Column {data-height=700}
-----------------------------------------------------------------------
### Coastal habitats diversity calculated with Shannon index (H)

```{r Coastal habitat diversity, fig.height=10}
# habitat diversity per country
rich$Shannon_H<- round(rich$Shannon_H, 2)
div_map <- merge(world, filter(rich, Zone == "Coastal"), by = "Country")
lines<-sf::st_as_sf(div_map)
div_map<- dplyr::select(lines, Country, Area=Shape_Area, Richness, Shannon_H)
mapview(div_map, zcol = "Shannon_H")


```


Oceanic map
=======================================================================

Column {data-height=700}
-----------------------------------------------------------------------
### Oceanic habitats diversity calculated with Shannon index (H)

```{r Oceanic habitat diversity, fig.height=10}
# habitat diversity per country
div_map <- merge(world, filter(rich, Zone == "Oceanic"), by = "Country")
lines<-sf::st_as_sf(div_map)
div_map<- dplyr::select(lines, Country, Area=Shape_Area, Richness, Shannon_H)
mapview(div_map, zcol = "Shannon_H")

```



Contacts
=======================================================================

Column {data-height=700}
-----------------------------------------------------------------------
### Authors

Fabio Favoretto and Joy Kumagai

For information and bug reports please contact: 

favoretto.fabio@gmail.com
jkumagai96@gmail.com

All data in this dashboard comes from open source datasets, however these have rules for sharing them
please **do not share** any of this information if not stricly related to the Blue Paper 10 drafting. 

If you might want some more information about these data or how to cite them, 
you can contact the authors and they will provide the info you needed. 