---
title: "Dashboard_test"
resource_files:
- data/EEZ_land.shx
- data/EEZ_land.prj
- data/EEZ_land.dbf
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    source_code: embed
    theme: spacelab
    vertical_layout: scroll
---

```{r global, include=FALSE}
library(flexdashboard)
library(datasets)
library(tidyverse)
library(stringr)
library(mapview)
library(maptools)
library(raster)
library(rgeos)
library(DT)
library(vegan)
library(plotly)
library(shiny)
library(ggthemes)
world <- shapefile("data/EEZ_land.shp", verbose = FALSE)
habitats <- readRDS("data/habitats.RDS")
```

Habitats dataset
=======================================================================

Column {data-height=700}
-----------------------------------------------------------------------

### All habitat data were intersected with Exclusive Economical Zone layer to calculate the area in ```km^2```, then were divided in coastal habitats and oceanic habitats. 
```{r}
datatable(
  habitats %>%
    select(Country, Zone, Habitat, Area),
  filter = 'top',
  rownames = FALSE
) %>%
  formatRound("Area", digits = 1)
```

Column {data-height=700}
-----------------------------------------------------------------------

### Richnes values are totals per each zone: Oceanic and Coastal, same for Shannon H: use filter or search to show different Countries and the next button below to scroll down the rows.
```{r Datatable diversity}
habitats %>%
  add_count(Country, Zone) %>%
  group_by(Country, Zone) %>%
  mutate(H = diversity(Area, index = "shannon", 2))  %>%
  select(Country, Zone, Richness = n, Shannon_H = H) %>%
  unique(.) -> rich

datatable(rich,
          filter = 'top',
          rownames = FALSE) %>%
  formatRound("Shannon_H", digits = 1)
```


Habitats
=======================================================================


Column {.sidebar}
-----------------------------------------------------------------------

This allow to select specific habitat to show their habitat area contribution

```{r}
selectInput(
      inputId = "Habitat",
      label = NULL,
      choices = unique(habitats$Habitat)
      )
```


Column
-----------------------------------------------------------------------

### Top 20 countries for habitat area

```{r}
dfInput <- reactive({
        habitats %>% filter(Habitat == input$Habitat)
    })
    
renderPlot({
        df1 <- dfInput() %>% arrange(Country, Area) %>% 
            top_n(20)
        ggplot(df1) +
            aes(x = reorder(Country, Area), y = Area) +
            geom_col(fill = "black") +
            labs(x = "", y = bquote("Area (" * m ^ 2 ~ ")")) +
            theme_economist_white() +
            theme(axis.text.x = element_text(angle = 90))+
            coord_flip()
    })
```



Coastal habitats diversity map
=======================================================================

Column {data-height=700}
-----------------------------------------------------------------------
### Coastal habitats diversity calculated with Shannon index (H)

```{r Coastal habitat diversity, fig.height=10}
# habitat diversity per country

habitats %>%
  group_by(Country, Zone) %>%
  summarise(H = diversity(Area, index = "shannon", 2)) -> rich

div_map <- merge(world, filter(rich, Zone == "Coastal"), by = "Country")

mapview(div_map, zcol = "H")

```


Oceanic habitats diversity map
=======================================================================

Column {data-height=700}
-----------------------------------------------------------------------
### Oceanic habitats diversity calculated with Shannon index (H)

```{r Oceanic habitat diversity, fig.height=10}
# habitat diversity per country
habitats %>%
  group_by(Country, Zone) %>%
  summarise(H = diversity(Area, index = "shannon", 2)) -> rich

div_map <- merge(world, filter(rich, Zone == "Oceanic"), by = "Country")
library(mapview)

mapview(div_map, zcol = "H")

```
